---
title: "Exponential distribution and CLT"
author: "Simon Lidberg"
date: "28th of August 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(knitr)
require(ggplot2)
```

## Overview  
This project will test the central limit theorem (CLT) with the exponential function.

## Simulations  
The default values for the simulations are stated below in the R code. The simulations are run 1000 times, each generating 40 rexp values with a lambda value of 0.2. The rexp values are either summarized with mean or var.

```{r simulations}
# Since we are using random numbers, we should set the seed value for reproducability.
set.seed(42)

# Lambda should be 0.2 for all simulations.
lambda <- 0.2

# The number of exponentials to average.
size <- 40

# The number of simulations to perform.
nrSimulations <- 1000

# The simulated means.
means <- replicate(nrSimulations, mean(rexp(size, lambda)))

# The simulated standard deviations.
vars <- replicate(nrSimulations, sd(rexp(size, lambda)))

# Print it nicely with kable.
Statistic <- c("Mean","Std. Deviation")
Sample <- c(round(mean(means),2),round(mean(vars),2))
Theoretical <- c(1/lambda, 1/lambda)
theTable <- data.frame(Statistic,Sample,Theoretical)
rownames(theTable) <- NULL
kable(theTable)

```

## Distribution  
Showing the histogram for the means, it resembles the density of the normal distribution, superimposed on the graph.

```{r}
hist(means, freq=FALSE, xlim= 5 +c(-3, 3), ylim = c(0, 0.7), xlab = "Means of rexp(40, 1/lambda)", main = "Histogram of the simulated values")
curve(dnorm(x, mean = 5, sd = 25/40), add = TRUE)
```

We can also use a Q-Q plot to show how closely it correlates to the normal distribution.

```{r}
qqnorm(means)
qqline(means)
```

The closer the points are to the straight line, the better the estimate is.
This is possible through the CLT.

We can conclude that the distribution is approximately normal and that the distribution of means of 40 exponentials behave as predicted by the CLT.