---
title: "Basic Inferential Data Analysis"
author: "Simon Lidberg"
date: "28th of August 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(datasets)
require(pastecs)
require(ggplot2)
require(dplyr)
```

## Overview
This analysis will compare the effect of Vitamin C on tooth growth in Guinea Pigs.

## Basic exploratory analysis  
From the help file we can see that the dataset consists of 60 observations with 3 variables.
The tooth length, the supplement type and the dose.

```{r LoadAndExplore}
# Load the dataset.
data <- ToothGrowth

# Show a summary of the dataset.
summary(data)

# Some summary statistics with grouping.
data %>% group_by(supp) %>% summarize(mean = mean(len), sd = sd(len))

# Show detailed statistics for len.
stat.desc(data[,1], norm = TRUE)

# Make an exploratory plot.
qplot(data = ToothGrowth, x = len, y = dose, facets = supp ~ .)
```
The exploratory analysis seems to indicate that VC combined with the largest dose results in the highest length of teeth and that OJ supplement form gives the higher mean length. Let's find out if that holds up with hypothesis testing and confidence intervals.

## Compare tooth growth by supp and dose.
Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)




## Conclusions
State your conclusions and the assumptions needed for your conclusions.
